---
title: "1. Getting Started with AMPH Forecast Suite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with AMPH Forecast Suite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The AMPH Forecast Suite provides a comprehensive toolkit for building infectious disease forecasting pipelines and hubs. This vignette demonstrates the basic workflow for using this package. This package and suite are still under active development, and we welcome contributions and feedback from the community.


## Installation

First, ensure you have R (version 4.0 or higher) and RStudio installed.

```{r}
# Check R version
R.Version()$version.string
```

Next, set your git credentials, if you have not already done this.

See this [guide](https://happygitwithr.com/https-pat) or [github website](information.https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/managing-your-personal-access-tokens) for more information on creating a GitHub personal access token (PAT).

```{r eval=FALSE}
# Set your git credentials (replace with your actual info)
install.packages("gitcreds")

gitcreds::gitcreds_set(
  username = "your_github_username",
  password = "your_github_token"
)
```


Next, install the AMPH Forecast Suite from CRAN or GitHub:

```{r eval=FALSE}
# Install remotes if you haven't already
install.packages("remotes")

# Install AMPH Forecast Suite
remotes::install_github("ACCIDDA/AMPH_Forecast_Suite")
```


Next, install the required forecasting packages:

```{r eval=FALSE}
library(AMPHForecastSuite)

# Install all forecasting packages
install_forecast_packages()
```

You can also selectively install specific categories:

```{r eval=FALSE}
# Install only hubverse and forecasting packages
install_forecast_packages(
  install_hubverse = TRUE,
  install_forecasting = TRUE,
  install_epi = FALSE,
  install_data = FALSE
)
```

## Setting Up Directory Structure
Set up a directory structure for your forecasting project:
```{r eval=FALSE}
# Create project directory and subdirectories
# this sets the working directory to the new project directory
setup_file_structure(project_dir = "My_Forecast_Project")

```



## Next steps

1. Define the forecasting effort:
  - Determine forecasting disease and targets
  - Identify data sources and geographies
2. Pull and manipulating data to use with forecasting models
3. Run forecasting models
4. Ensemble, visualize, and evaluate forecasts
5. Share forecasts with stakeholders


For more information, see the package documentation and function help pages.
